Adversarial Training of Invariant Features for Speech Recognition

### 2017-03-22 interspeech/paper.tex
Change citation style

\usepackage{natbib}
\usepackage{INTERSPEECH_v2}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[draft]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

@@

    Several model and feature based adaptation methods such as Maximum 
    Likelihood Linear Regression (MLLR),  feature-based MLLR and 
    i-vectors~\citep{saon2013speaker} have been proposed to handle speaker 
    variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle 
    environment variability.
    % The end-to-end part has become sort of irrelevant given that the
    % experiments are actually with hybrid systems and this version of the paper will not
    % have to be sold as being end-to-end anymore.
    With the increasing success of Deep Neural 
    Network~(DNN) acoustic models for ASR investigated 
    in~\citep{hinton2012deep,seide2011conversational,sainath2011making}, 
    and in more recent works~\citep{miao2015eesen,sainath2015learning} 
    complicated structure of acoustic conditions is modeled within a single network.
    This allows us to take 
    advantage of the network's ability to learn highly non-linear feature 
    transformations, with greater flexibility in constructing training 
    objective functions that promote learning of noise invariant 
    representations.
    At the same time, the more complex relation between the parameters of deep
    neural networks and the functions or distributions they represent, makes it
    impossible to simply use the same types of adaptation and robustness methods
    that have been developed for GMM-based systems.

    The main idea of this work is to force the acoustic model 
    to learn a representation which is invariant to noise conditions, instead of 
    explicitly using noise robust acoustic features 
    (Section~\ref{sec:invariant-speech}). This type of noise-invariant 
    training requires noise-condition labels during training only. It is 
    related to the idea of generative adversarial networks (GAN) and the 
    gradient reverse method proposed by~\cite{goodfellow2014generative} 
    and~\cite{ganin2014unsupervised} respectively. 
    First, we discuss these two methods in Section~\ref{sec:background}.
    Then, we discuss previous work in Section~\ref{sec:relatedwork}. 
    Finally, we present results on the Aurora-4 speech 
    recognition task in Section~\ref{sec:experiments} and summarize 
    our findings in Section~\ref{sec:discussion}.

@@

    Many neural network speaker adaptation methods are based on
    i-vectors. The most common way to exploit i-vectors is by providing them as additional
    inputs~\citep{senior2014improving,saon2013speaker}. This means that i-vectors also need to
    be available during testing and this may not always be practical.

    In the specific case of noise robustness, one can use speech enhancement to
    obtain more robust features and neural networks have been used for this for
    decades~\citep{knecht1995neural}. A further step in this direction is to train
    systems to recognize speech and perform speech enhancement jointly~\citep{narayanan2014joint}. 
    We argue that speech enhancement is a more difficult task to learn than
    invariance to certain noise conditions and more likely to suffer from
    overfitting when the amount of available noisy data is limited.
    Another downside of enhancement approaches is that there is no reason to
    expect them to generalize well to noise conditions that were not available during
    training.

    Another adaptation strategy is to retrain the model parameters using a very small
    number of utterances from the new domain while making sure that the model
    doesn't stray too far away from the parameter values that were obtained from
    the train data~\citep{yu2013kl}.
    The adaptation to new domains can also be kept under control by limiting
    the number of parameters which are adapted to the new data or by limiting the
    adaptation to a rescaling of the hidden unit activations~\citep{swietojanski2014learning}.
    The most important difference between these approaches and our work is that
    we don't adapt the model parameters at all after training. In many
    situations this may not be possible and retraining of models may require more
    from the hardware on which the speech recognition system is implemented.

    Recently, in a work by~\cite{yusuke2016adversarial} a multi-layer sigmoidal network was trained 
    in an adversarial fashion on an in-house transcription task corrupted by noise.
    This work is very similar to our approach but we evaluate our methods on
    more challenging benchmark while investigating different numbers of noise
    conditions. Our work also differs due to the use of more modern rectifier
    activation based classification networks.
    Finally, in other very recent work~\citep{saon2017english} a form of adversarial training with a network
    that predicts an i-vector using the mean square error loss was used.

@@

\section{Experiments}
\label{sec:experiments}
We experimentally evaluated our approach   
on the well-benchmarked Aurora-4 \citep{parihar2002aurora} noisy speech recognition task. Aurora-4
is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
six categories which were added to the clean data. Every clean and noisy
utterance has been 
filtered to simulate the phone quality recording using P.341 \citep{pariharsignal}. The training
data contains 4400 clean utterances and 446 utterances for each noise condition,
i.e., a total of 2676 noisy utterances.
The test set consists of clean data, data corrupted by 6 noise types, and data 
recorded with a different microphone for both the clean and noisy conditions.

For both the clean and noisy data, we extracted 40-dimensional Mel-filterbank features with their deltas and 
delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
features that were subsequently mean and variance normalized. The baseline
acoustic model was a 6-layer 
DNN with 2048 rectified linear units at every layer. It was trained using 
momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
annealing~\citep[the learning rate is halved if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.

@@

\section{Conclusion}
\label{sec:discussion}
    This paper presents the application of generative adversarial networks and 
    invariance training for noise robust speech recognition. We show that invariance training 
    helps the ASR system to generalize better to unseen noise conditions and improves 
    the word error rate when a small number of noise types are seen during training. Our 
    experiments show that in contrast to the image recognition task, in speech 
    recognition, the task of the domain classification ($D$) network is harder. Therefore, the 
    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 
    research includes enhancements to the domain adaptation network while exploring 
    alternative network architectures and invariance-promoting loss functions.

@@

We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
developers of Theano~\citep{theano2016theano}, Blocks, and Fuel~\citep{MerrienboerBDSW15} 
for great toolkits. 
The authors would like to acknowledge the support of the following agencies for
research funding and computing support: NSERC, Calcul Qu\'{e}bec, Compute Canada,
the Canada Research Chairs and CIFAR. Dmitriy Serdyuk thanks Samsung for partial
financial support of this work and IBM Watson where he performed this work 
during an internship.

\bibliographystyle{IEEEtranN}

>>>>

\usepackage{INTERSPEECH_v2}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[draft]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

@@

    Several model and feature based adaptation methods such as Maximum 
    Likelihood Linear Regression (MLLR),  feature-based MLLR and 
    i-vectors~\cite{saon2013speaker} have been proposed to handle speaker 
    variability; and Noise Adaptive Training~(NAT;~\cite{kalinli2010noise})
    and Vector Taylor Series~(VTS;~\cite{un1998speech}) to handle 
    environment variability.
    % The end-to-end part has become sort of irrelevant given that the
    % experiments are actually with hybrid systems and this version of the paper will not
    % have to be sold as being end-to-end anymore.
    With the increasing success of Deep Neural 
    Network~(DNN) acoustic models for ASR investigated 
    in~\cite{hinton2012deep}\cite{seide2011conversational}\cite{sainath2011making}, 
    and in more recent works~\cite{miao2015eesen}\cite{sainath2015learning} 
    complicated structure of acoustic conditions is modeled within a single network.
    This allows us to take 
    advantage of the network's ability to learn highly non-linear feature 
    transformations, with greater flexibility in constructing training 
    objective functions that promote learning of noise invariant 
    representations.
    At the same time, the more complex relation between the parameters of deep
    neural networks and the functions or distributions they represent, makes it
    impossible to simply use the same types of adaptation and robustness methods
    that have been developed for GMM-based systems.

    This work is a continuation of our preliminary results presented as a workshop 
    paper~\cite{serdyuk2016invariant}.
    The main idea of this work is to force the acoustic model 
    to learn a representation which is invariant to noise conditions, instead of 
    explicitly using noise robust acoustic features 
    (Section~\ref{sec:invariant-speech}). This type of noise-invariant 
    training requires noise-condition labels during training only. It is 
    related to the idea of generative adversarial networks (GAN) and the 
    gradient reverse method proposed by~\cite{goodfellow2014generative} 
    and~\cite{ganin2014unsupervised} respectively. 
    First, we discuss these two methods in Section~\ref{sec:background}.
    Then, we discuss previous work in Section~\ref{sec:relatedwork}. 
    Finally, we present results on the Aurora-4 speech 
    recognition task in Section~\ref{sec:experiments} and summarize 
    our findings in Section~\ref{sec:discussion}.

@@

    Many neural network speaker adaptation methods are based on
    i-vectors. The most common way to exploit i-vectors is by providing them as additional
    inputs~\cite{senior2014improving,saon2013speaker}. This means that i-vectors also need to
    be available during testing and this may not always be practical.

    In the specific case of noise robustness, one can use speech enhancement to
    obtain more robust features and neural networks have been used for this for
    decades~\cite{knecht1995neural}. A further step in this direction is to train
    systems to recognize speech and perform speech enhancement jointly~\cite{narayanan2014joint}. 
    We argue that speech enhancement is a more difficult task to learn than
    invariance to certain noise conditions and more likely to suffer from
    overfitting when the amount of available noisy data is limited.
    Another downside of enhancement approaches is that there is no reason to
    expect them to generalize well to noise conditions that were not available during
    training.

    Another adaptation strategy is to retrain the model parameters using a very small
    number of utterances from the new domain while making sure that the model
    doesn't stray too far away from the parameter values that were obtained from
    the train data~\cite{yu2013kl}.
    The adaptation to new domains can also be kept under control by limiting
    the number of parameters which are adapted to the new data or by limiting the
    adaptation to a rescaling of the hidden unit activations~\cite{swietojanski2014learning}.
    The most important difference between these approaches and our work is that
    we don't adapt the model parameters at all after training. In many
    situations this may not be possible and retraining of models may require more
    from the hardware on which the speech recognition system is implemented.

    Recently, in a work by~\cite{yusuke2016adversarial} a multi-layer sigmoidal network was trained 
    in an adversarial fashion on an in-house transcription task corrupted by noise.
    This work is very similar to our approach but we evaluate our methods on
    more challenging benchmark while investigating different numbers of noise
    conditions. Our work also differs due to the use of more modern rectifier
    activation based classification networks.
    Finally, in other very recent work~\cite{saon2017english} a form of adversarial training with a network
    that predicts an i-vector using the mean square error loss was used.

@@

\section{Experiments}
\label{sec:experiments}
We experimentally evaluated our approach   
on the well-benchmarked Aurora-4 \cite{parihar2002aurora} noisy speech recognition task. Aurora-4
is based on the Wall Street Journal corpus (WSJ0). It contains noises of 
six categories which were added to the clean data. Every clean and noisy
utterance has been 
filtered to simulate the phone quality recording using P.341 \cite{pariharsignal}. The training
data contains 4400 clean utterances and 446 utterances for each noise condition,
i.e., a total of 2676 noisy utterances.
The test set consists of clean data, data corrupted by 6 noise types, and data 
recorded with a different microphone for both the clean and noisy conditions.

For both the clean and noisy data, we extracted 40-dimensional Mel-filterbank features with their deltas and 
delta-deltas spliced over $\pm$5 frames, resulting in 1320 input 
features that were subsequently mean and variance normalized. The baseline
acoustic model was a 6-layer 
DNN with 2048 rectified linear units at every layer. It was trained using 
momentum-accelerated stochastic gradient descent for 15 epochs with new-bob 
annealing~\cite[the learning rate is halved if no improvement on the validation set, as in][]{morgan1995continuous,sainath2011making}.

@@

\section{Conclusion}
\label{sec:discussion}
    This paper presents the application of generative adversarial networks and 
    invariance training for noise robust speech recognition. We show that invariance training 
    helps the ASR system to generalize better to unseen noise conditions and improves 
    the word error rate when a small number of noise types are seen during training. Our 
    experiments show that, relative to the image recognition tasks reported
    by \cite{ganin2014unsupervised}, the task of the domain classification ($D$) network is
    harder for speech recognition. Therefore, the 
    gradient of the $L_3$ term in Eq.~\ref{eq:grm} is noisier. Future 
    research includes enhancements to the domain adaptation network while exploring 
    alternative network architectures and invariance-promoting loss functions.

@@

We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
developers of Theano~\cite{theano2016theano}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
for great toolkits. 
The authors would like to acknowledge the support of the following agencies for
research funding and computing support: NSERC, Calcul Qu\'{e}bec, Compute Canada,
the Canada Research Chairs and CIFAR. Dmitriy Serdyuk thanks Samsung for partial
financial support of this work and IBM Watson where he performed this work 
during an internship.

\bibliographystyle{IEEEtran}
